A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-15 ionic_liquids.tex
Further changes in dilute CO2 solutions

%% Group authors per affiliation:
\author[plapiqui,msk]{A.J.~Silveira}
\author[plapiqui]{S.~Pereda}
\author[eq-ufrj,coppe-ufrj]{F.W.~Tavares}
\author[eq-ufrj]{C.R.A.~Abreu\corref{cor1}}
\ead{abreu@eq.ufrj.br}

\address[plapiqui]{Planta Piloto de Ingenier\'ia Qu\'imica (PLAPIQUI), Universidad Nacional del Sur, Bah\'ia Blanca, Argentina}
\address[msk]{Computational \& Systems Biology Program, Memorial Sloan Kettering Cancer Center, New York, NY, USA}
\address[eq-ufrj]{Chemical Engineering Department, Escola de Qu\'imica, Universidade Federal do Rio de Janeiro, Rio de Janeiro, Brazil}
\address[coppe-ufrj]{Chemical Engineering Program, Alberto Luiz Coimbra Institute for Graduate Studies and Research in Engineering (COPPE), Universidade Federal do Rio de Janeiro, Rio de Janeiro, Brazil}

\cortext[cor1]{Corresponding author}

@@

\begin{keyword}
ionic liquids \sep solvation free energy \sep  \ce{CO_2} Henry constant \sep infinite-dilution activity coefficient 
\end{keyword}

@@

One can see in Table \ref{table:henry} that the only model able to reproduce the experimental Henry's law constant of \ce{CO_2} in \ce{[emim][B(CN)_4]} with small deviation is that of Koller \textit{et al}. \cite{Koller_2012}.
The worst results are obtained with the models of Batista \textit{et al}. \cite{Batista_2015} and Liu \textit{et al}. \cite{Liu_2014}.
These models fail in part because of the small values they predict for $\Delta G_\text{\tiny vdW}$.
In order to analyze the model distinctions in more detail, in Fig.~\ref{fig:deltag} we plot the profiles of $\Delta G_\text{\tiny vdW}$ as a function of the coupling parameter $\lambda_\text{\tiny vdW}$.

If we look at the $\Delta G_\text{\tiny vdW}$ profiles in Fig.~\ref{fig:deltag}, we observe that for the models of Batista \textit{et al}. \cite{Batista_2015} and Liu \textit{et al}.~\cite{Liu_2014} the maximum occurs at lower values of energy, suggesting that these models underestimate the atomic interactions, which may also explain their failure to reproduce the experimental density.

@@

The $\Delta G_{\, \text{sim}}$ values for the models of Koller \textit{et al}. \cite{Koller_2012} and K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} suggest a more favorable solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}. 
The spatial distribution functions (SDFs) \cite{Svishchev_1993} shown in Figs.~\ref{fig:sdf_ions} confirm this result.
The SDFs of the ions around a \ce{CO_2} molecule were obtained by using the software packages TRAVIS \cite{Brehm_2011} and VMD \cite{HUMP96}.
For that, we carried out NPT-MD simulations of \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-\ce{[emim][NTf_2]} systems with a \ce{CO_2} molar fraction of 0.1.
The isovalues were visually determined so as to observe the first solvation shell.
If we compare Figs.~\ref{fig:sdf_ions} (a) and (b), one notes that the region with a low probability of finding an ion is larger for \ce{[emim][NTf_2]}.
Also, the first solvation shell due to \ce{[B(CN)_4^-} occurs at a slightly shorter distance than for \ce{[NTf_2]^-}, as one sees from Fig.~\ref{fig:sdf_ions}(c).

>>>>

%% Group authors per affiliation:
\author[plapiqui]{A.J.~Silveira\corref{curradd}}
\author[plapiqui]{S.~Pereda}
\author[eq-ufrj,coppe-ufrj]{F.W.~Tavares}
\author[eq-ufrj]{C.R.A.~Abreu\corref{cor1}}
\ead{abreu@eq.ufrj.br}

\address[plapiqui]{Planta Piloto de Ingenier\'ia Qu\'imica (PLAPIQUI), Universidad Nacional del Sur, Bah\'ia Blanca, Argentina}
\address[eq-ufrj]{Chemical Engineering Department, Escola de Qu\'imica, Universidade Federal do Rio de Janeiro, Rio de Janeiro, Brazil}
\address[coppe-ufrj]{Chemical Engineering Program, Alberto Luiz Coimbra Institute for Graduate Studies and Research in Engineering (COPPE), Universidade Federal do Rio de Janeiro, Rio de Janeiro, Brazil}

\cortext[curradd]{Current address: Computational \& Systems Biology Program, Memorial Sloan Kettering Cancer Center, New York, NY, USA}
\cortext[cor1]{Corresponding author}

@@

\begin{keyword}
ionic liquids \sep solvation free energy \sep \ce{CO_2} Henry's law constant \sep infinite-dilution activity coefficient
\end{keyword}

@@

One can see in Table \ref{table:henry} that the only model able to reproduce the experimental Henry's law constant of \ce{CO_2} in \ce{[emim][B(CN)_4]} with small deviation is that of Koller \textit{et al}. \cite{Koller_2012}.
The worst results are obtained with the models of Batista \textit{et al}. \cite{Batista_2015} and Liu \textit{et al}. \cite{Liu_2014}.
These models fail in part because of the small values they predict for $\Delta G_\text{\tiny vdW}$.
In order to analyze the model distinctions in more detail, in Fig.~\ref{fig:deltag} we plot free energy profiles with respect to the coupling parameter $\lambda_\text{\tiny vdW}$, as obtained from the simulations with all \ce{[emim][B(CN)_4]} models.
We observe that the model of Batista \textit{et al}. \cite{Batista_2015} is the one that predicts the smallest free-energy barrier, immediately followed by the model of Liu \textit{et al}. \cite{Liu_2014}.
This observation suggests that these models underestimate the magnitude of the solvent-solvent interactions, thus making it easier to open a cavity in the system for inserting the solute.
Besides, it also explains why these models underestimate the density of \ce{[emim][B(CN)_4]}, such as seen in Table~\ref{table:props_dsf}.
The Liu \textit{et al}. \cite{Liu_2014} model predicts the smallest value of $\Delta G_\text{\tiny vdW}$ in spite of exhibiting only the second smallest free-energy barrier.
This occurs because the attractive van der Waals interactions compensate the cavity formation more favorably than in the case of Batista \textit{et al}. \cite{Batista_2015}, as one can observe in Fig.~\ref{fig:deltag} as $\lambda_\text{\tiny vdW}$ approaches unity.
Nevertheless, these distinctions end up being canceled out when the charges of the \ce{CO_2} molecule are introduced, since $\Delta G_\text{\tiny vdW}$ and $\Delta G_\text{\tiny Coul}$ sum up to very similar $\Delta G_\text{sim}$ results for both models.

@@

The $\Delta G_\text{sim}$ values for the models of Koller \textit{et al}. \cite{Koller_2012} and K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} suggest a more favorable solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}. 
The spatial distribution functions (SDFs) \cite{Svishchev_1993} shown in Figs.~\ref{fig:sdf_ions} confirm this result.
The SDFs of the ions around a \ce{CO_2} molecule were obtained by using the software packages TRAVIS \cite{Brehm_2011} and VMD \cite{HUMP96}.
For that, we carried out NPT-MD simulations of \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-\ce{[emim][NTf_2]} systems with a \ce{CO_2} molar fraction of 0.1.
The isovalues were visually determined so as to observe the first solvation shell.
If we compare Figs.~\ref{fig:sdf_ions} (a) and (b), one notes that the region with a low probability of finding an ion is larger for \ce{[emim][NTf_2]}.
Also, the first solvation shell due to \ce{[B(CN)_4^-} occurs at a slightly shorter distance than for \ce{[NTf_2]^-}, as one sees from Fig.~\ref{fig:sdf_ions}(c).
